### 2022-08-24 thesis/thesis.tex
apply aug 23 notes

\usepackage[b4paper,left=0.8in,right=0.8in,top=0.8in,bottom=0.8in]{geometry}
\usepackage{tikz-qtree}
\usepackage{algorithm}
\usepackage{algpseudocode}
\makeatletter
\renewcommand{\ALG@beginalgorithmic}{\footnotesize}
\makeatother
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{showkeys}
\usepackage{amsmath}

@@

\begin{lemma}[headProgress] \label{lem::headProgress}
\nf{n.head} is non-decreasing over time. If \nf{n.blocks[i]$\neq$null} and \nf{i$.$0} then \nf{n.blocks[i].end\sub{left}} $\geq$ \nf{n.blocks[i-1].end\sub{left}} and \nf{n.blocks[i].end\sub{right}} $\geq$ \nf{n.blocks[i-1].end\sub{right}}.
\end{lemma}
\begin{proof}
The first claim follows trivially from the pseudocode since \nf{n.head} is only incremented in the pseudocode in lines \ref{incrementHead1} and \ref{incrementHead2} of \nf{Refresh()} .

@@

%Suppose the previous true returning \nf{n.Refresh()} took place in time $(t_i^\prime-t_t^\prime)$. By induction on the number of succsuful \nf{Refresh()}'s \nf{n} we know that \nf{ops(EST\sub{n.left, t\sub{i}$^\prime$})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{i}$^\prime$})} $\in$ \nf{ops(EST\sub{n, t\sub{t}$^\prime$})}. Operations that are in \nf{ops(EST\sub{n.left, t\sub{i}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{i}})} $\in$ \nf{ops(EST\sub{n, t\sub{t}})} but not in \nf{ops(EST\sub{n.left, t\sub{i}$^\prime$})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{i}$^\prime$})} $\in$ \nf{ops(EST\sub{n, t\sub{t}$^\prime$})} are in the block returned by \nf{n.CreateBlock()} by lemma 10. 
%By Lemma~\ref{lem::createBlock} \nf{new} contains \nf{n}'s childrens' established blocks after Line \ref{readHead} which is  appended to \nf{n.blocks[]} by \nf{TryAppend} in Line 313.
\end{proof}
\begin{figure}[hbt]
  \center\includegraphics[width=5in]{pics/trueRefresh}
  \caption{New established operations of the left child are in the new block.}
\end{figure}
\begin{lemma}[Stronger True Refresh] \label{lem::prectrueRefresh}
  Let $t_i$ be the time an instance of \nf{n.Refresh()} read the head (Line \ref{readHead}) and $t_t$ be the time its \nf{TryAppend(new, s)} terminates with  and returns \nf{true} (Line \ref{cas}). We have \nf{ops(EST\sub{n.left, t\sub{i}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{i}})} $\subseteq$ \nf{ops(n.blocks)}.
\end{lemma}
%\it{Mention this new block is establish in t\sub{t}. last sentece is not complete}

@@

\begin{lemma}[Double Refresh] \label{doubleRefresh}
  Consider two consecutive instances $R_1$, $R_2$ of \nf{Refresh()} on \nf{n} by process $p$. Let $t_1$ be the time $R_1$ is invoked and $t_2$ be the time $R_2$ terminated. If $R_1$ and $R_2$ both fail and return false, then we have \nf{ops(EST\sub{n.left, t\sub{1}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{1}})} $\subseteq$ \nf{ops(EST\sub{n, t\sub{2}})}.
\end{lemma}
\begin{proof}

If Line \ref{cas} of $R_1$ or $R_2$ returns \nf{true}, then the claim is held by Lemma~\ref{lem::trueRefresh}. Let $R_1$ read $i$ and $R_2$ read $i+1$ from Line \ref{readHead}. If $R_2$ reads some value greater than $i+1$ in Line \ref{readHead} it means a successful instance of \nf{Refresh()} started after Line \ref{readHead} of $R_1$ and finished its Line \ref{incrementHead1} or \ref{incrementHead2} before \ref{readHead} of $R_2$, from Lemma~\ref{lem::trueRefresh} by the end of this instance \nf{ops(EST\sub{n.left, t\sub{1}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{1}})} has been propagated.

Since $R_2$'s \nf{TryAppend()} returns \nf{false} then there is another successful instance $R_2^\prime$ of \nf{n.Refresh()} that has done \nf{TryAppend()} successfully into \nf{n.blocks[i+1]} before $R_2$ tries to append. Since $R_2^\prime$ creates the block after reading the value $i+1$ from \nf{n.head}(Line \ref{readHead}) and $R_1$ reads the value $i$ from \nf{n.head} and the \nf{head}'s value is increaing by Lemma \ref{lem::headProgress} then $t_{R2^\prime \; \ref{readHead}} > t_{R1 \; \ref{readHead}} > t_1$ (See Figure \ref{fig::doubleRefresh}). By Lemma \ref{lem::prectrueRefresh} after $R_2^\prime$'s \nf{CAS} we have \nf{ops(EST\sub{n.left, t\sub{1}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{1}})} $\subseteq$ \nf{ops(n.blocks)}. Also by Lemma \ref{lem::headInc} on $R_2$ the value  of \nf{n.head}  head is more than $i+1$ after $R_2$\' terminates, so the block appended by $R_2^\prime$ to \nf{n} is established by then (\nf{n.head>=i+2>i+1}). To summarize $t_1$ is before $R_2^\prime$'s \nf{read n.head} and $R_2^\prime$'s successful \nf{CAS} is before $R_2$'s termination. So by Lemma \ref{lem::prectrueRefresh} \nf{ops(EST\sub{n.left, t\sub{1}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{1}})} $\subseteq$ \nf{ops(EST\sub{n, t\sub{2}})}.
%It is obvious that the \nf{new} constructed by the second Refresh in Line 36 contains the blocks in \nf{new} by $R_1$ which $R_1^\prime$ did not contain, since \nf{n.head} is only increasing (Maybe Lemma~\ref{lem::oldnewOrder}). If $R_2$ succeeds by Lemma~\ref{lem::trueRefresh} the claim holds. If not, it is deduced that \nf{n.blocks[h]} was not \nf{null} before $R_2$'s \nf{CAS}. Furthermore, \nf{n.blocks[h]} was \nf{null} before reading \nf{h} by $R_1$. So there is a successful \nf{Refresh()} after the read of \nf{h} in $R_1$ and before the \nf{CAS} of $R_2$. This \nf{Refresh()} contains all the new established operations before Line 35 (Maybe Lemma~\ref{lem::oldnewOrder}) and by Lemma~\ref{lem::trueRefresh} our claim holds. See Figure~\ref{}.
\end{proof}

\begin{figure}[hbt]
  \center\includegraphics[width=6in]{pics/compactdouble.png}
  \caption{\label{fig::doubleRefresh}$t1<$ $r_1$ reading head $<$ incrementing \nf{n.head} from $i$ to $i+1$ $<$ $R_2^\prime$ reading head $<$ \nf{TryAppend(i+1)} $<$ incrementing \nf{n.head} from $i+1$ to $i+2$ $<t2$} 
\end{figure}
\it{this chain with more depth should be in the proof}

@@

\begin{definition}
  $t_{before \; line}$ is the immediate time before running Line $line$.   $t_{before \; line}$ is the immediate time after running Line $line$.
\end{definition} 
\begin{corollary}
  \nf{ops(EST\sub{n.left, t\sub{before \ref{firstRefresh}}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{before \ref{firstRefresh}}})} $\subseteq$ \nf{ops(EST\sub{n, t\sub{after \ref{secondRefresh}})}}
\end{corollary}
\begin{proof}
  If the first \nf{Refresh()} in line \ref{firstRefresh} returns \nf{true} then by Lemma \ref{lem::trueRefresh} the claim holds. Also if first \nf{Refresh()} failed and the second \nf{Refresh()} succeeded the claim still holds by Lemma \ref{lem::trueRefresh}. Finally, if both failed the claim is satisfied by Lemma \ref{doubleRefresh}.
\end{proof}

@@

\begin{lemma}[Get correctness] \label{get}
If \nf{n.blocks[b].num\sub{enq}$\geq$i} then \nf{n.GetENQ(b,i)} returns the \nf{element} enqueued by $E_i(n,b)$.
\end{lemma}
\begin{proof}
We are going to prove this lemma by induction on the height of node \nf{n}. For the base case \nf{n} is a leaf. Leaf blocks each contain exactly one operation, so by the preconditions of \nf{GetENQ()}, only \nf{n.GetENQ(b,1)} can be called and \nf{n.blocks[b]} contains an enqueue. At Line \ref{getBaseCase} \nf{n.GetENQ(b,1)} returns the \nf{element} of the \nf{enqueue} operation stored in the $b$th block of leaf \nf{n} .

For the induction step we prove \nf{n.GetENQ(b,i)} returns $E_i(n,b)$, if \nf{n.child.GetENQ(b,i)} returns $E_i(n.child,b)$. In  Line \ref{leftOrRight} it is decided for the non-leaf nodes that the \nf{i}th enqueue in  \nf{b}th block  of internal node \nf{n} is in the \nf{n.blocks[b]}'s left child or right child subblocks. From Definition  \ref{ordering} of $E(n,b)$ we know enqueue operations in a block are ordered by their process id and since the leaves of the tree are ordered by process id from left to right, thus operations from the left subblocks come before operations from the right subblocks in a block (See Figure \ref{figGet}). Furthermore the \nf{num\sub{enq-left}} field  in a block stores the number of \nf{enqueue()} operations from the blocks's subblocks in the left child of \nf{n}. So $i$th enqueue operation is propagated from the right child if \nf{i} is greater than \nf{b.num\sub{enq-left}}. otherwise we should search for the $i$th enqueue in the left child. By definition \ref{def::ops} and \ref{def::subblock} we need to search in subblocks of \nf{n.blocks[b]} from the range \texttt{n.left.blocks[n.blocks[i-1].end\textsubscript{left}+1..n.blocks[i].end\textsubscript{left}]} $\cup$ \texttt{n.right.blocks[n.blocks[i-1].end\textsubscript{right}+1..n.blocks[i].end\textsubscript{right}]}.

If the $i$th \nf{enqueue} of \nf{n.blocks[b]} is in the left child it would be $i$th enqueue in \texttt{n.left.blocks[n.blocks[i-1].end\textsubscript{left}+1..n.blocks[i].end\textsubscript{left}]} by Definition \ref{def::subblock}. Also we know there are $eb=n.blocks[b-1].sum_{enq-left}$ enqueues in the blocks before this range, so $E_i(n,b)$ is $E_{i+eb}(n.left)$ which is $E_{i^\prime}(n.left,b^\prime)$ for some $b^\prime$ and $i^\prime$. We can compute $b^\prime$ search for $i+eb$th enqueue in \tt{n.left} and $i^\prime$ is \tt{i+eb-n.left.blocks[}$b^\prime-1$\tt{].sum\sub{enq}}. The parameters in \ref{leftChildGet} are for searching $E_{i+eb}(n.left)$ in \nf{n.left.block} in the expected range of blocks, so this \nf{BSearch} returns the index of the subblock containing $E_i(n,b)$.

Else if the enqueue we are looking for is in the right child then there are \nf{n.blocks[b].num\sub{enq-left}} enqueues ahead of it in \nf{n.blocks[b]} but not in \texttt{n.right.blocks[n.blocks[i-1].end\textsubscript{right}+1..n.blocks[i].end\textsubscript{right}]}. So we need to search for \nf{i-n.blocks[b].num\sub{enq-left}+ n.blocks[b-1].sum\sub{enq-right}} (Line \ref{rightChildGet}). Other parameters are assigned similar for the left child. So in both cases the direct subblock containing $E_i(n,b)$ is computed in Lines \ref{leftChildGet} and \ref{rightChildGet}.

 Finally, \nf{n.child.GetENQ()} is invoked on the subblock containing $E_i(n,b)$ which returns $E_i(n,b)$ by the hypothesis of the induction.
\end{proof}

@@

\begin{lemma}[DSearch correctness] \label{dsearch}
Assume \nf{root.blocks[end].sum\sub{enq}}$\geq$\nf{e} and $E_e(root)$'s \nf{element} is the response to some \nf{Dequeue()} operation in \nf{root.blocks[end]}.  \nf{DSearch(e, end)} returns \nf{<b, i>} such that $E_i(root,b)= E_e(root)$.
\end{lemma}
\begin{proof}
It is trivial to see that the doubling search from \nf{root.blocks[end]} to \nf{root.blocks[0]} will find $E_e(root)$ eventually. Because \nf{root.blocks[].sum\sub{enq}} is an increasing value from 0 to some value greater than \nf{e}. So there is a \nf{b} that \nf{root.blocks[b].sum\sub{enq}}$>e$ but \nf{root.blocks[b-1].sum\sub{enq}}$<e$.

First we show \nf{end-b}$\leq 2 \times$\nf{root.blocks[b].\size+root.blocks[end].\size+1}. From line \ref{addOP}, we know that size of the every block in the tree is greater than 0. So each block in \nf{root.blocks[b..end]} contains at least one \nf{Enqueue} or at least one \nf{Dequeue}. Suppose there were more than \nf{root.blocks[b].\size}\nf{Dequeue}s in \nf{root.blocks[b+1..end-1]}. Then the  queue would become empty at some point after \nf{blocks[b]}'s last operations and before \nf{root.blocks[end]}'s first operation. Which means the response to to a \nf{Dequeue} in \nf{root.blocks[end]} could not be in $E(n,b)$. Furthermore since the size of the queue would become \nf{root.blocks[end].\size} after the \nf{root.blocks[end]}, there cannot be more than \nf{root.blocks[b].\size + root.blocks[end].\size} \nf{Enqueue}s. Because there can be at most \nf{root.blocks[b].\size}\nf{Dequeue}s and the final size is \nf{root.blocks[end].\size}. Overall there can be at most $2 \times$\nf{root.blocks[b].\size+ root.blocks[end].\size} operations in \nf{root.blocks[b+1..end-1]} and since each block size is $\geq 1$ thus there are at most $2 \times$\nf{root.blocks[b].\size+ root.blocks[end].\size} blocks in between \nf{root.blocks[b]} and \nf{root.blocks[end]}. So \nf{end-b}$\leq 2 \times$\nf{root.blocks[b].\size+root.blocks[end].\size+1}. See Figure \ref{end-b}.

@@

\begin{lemma}[Index correctness]
 If \nf{n.blocks[b].num\sub{deq}}$\geq$\nf{i} then \nf{n.IndexDEQ(b,i)} returns the rank in $D(root)$ of $D_{n,b,i}$.
\end{lemma}
\begin{proof}
We will prove this by induction on the distance of \nf{n} from the \nf{root}. We can see the base case where \nf{n} is root is trivial (Line \ref{indexBaseCase}).
  In the non-root nodes \nf{n.IndexDEQ(b,i)} computes the superblock of the $i$th Dequeue in the $b$th block of \nf{n} in \nf{n.parent} by Lemma \ref{superBlock} (Line \ref{computeSuper}). After that the order in $D(n.parent, superblock)$ is computed. Note that by Lemma \ref{blockSize} in each block there is at most one operation from each process and operations of one type are ordered based on the order in the subblocks (See Figure \ref{fig::orderFromSubblocks}). Finally \nf{index()} is called on \nf{n.parent} recursively and it returns the correct response from induction hypothesis. If the operation was propagated from the right child the number of dequeues from the left child are added to it (Line \ref{considerRight}), because the left child operations come before the right child operations (Definition \ref{ordering}).
\end{proof}
\textit{Make sure to show preconditions of all invocation of \nf{BSearch} are satisfied.}

>>>>

\usepackage[b4paper,left=0.8in,right=0.8in,top=0.8in,bottom=0.8in]{geometry}
\usepackage{tikz-qtree}
\usepackage{algorithm}
\usepackage{algpseudocode}
\makeatletter
\renewcommand{\ALG@beginalgorithmic}{\footnotesize}
\makeatother
\usepackage{graphicx}
\usepackage{subcaption}
%\usepackage{showkeys}
\usepackage{amsmath}

@@

\begin{lemma}[headProgress] \label{lem::headProgress}
\nf{n.head} is non-decreasing over time. If \nf{n.blocks[i]$\neq$null} and \nf{i$.$0} then $\nf{n.blocks[i].end\sub{left}} \geq \nf{n.blocks[i-1].end\sub{left}}$ and $\nf{n.blocks[i].end\sub{right}} \geq \nf{n.blocks[i-1].end\sub{right}}$.
\end{lemma}
\begin{proof}
The first claim follows trivially from the pseudocode since \nf{n.head} is only incremented in the pseudocode in lines \ref{incrementHead1} and \ref{incrementHead2} of \nf{Refresh()} .

@@

%Suppose the previous true returning \nf{n.Refresh()} took place in time $(t_i^\prime-t_t^\prime)$. By induction on the number of succsuful \nf{Refresh()}'s \nf{n} we know that \nf{ops(EST\sub{n.left, t\sub{i}$^\prime$})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{i}$^\prime$})} $\in$ \nf{ops(EST\sub{n, t\sub{t}$^\prime$})}. Operations that are in \nf{ops(EST\sub{n.left, t\sub{i}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{i}})} $\in$ \nf{ops(EST\sub{n, t\sub{t}})} but not in \nf{ops(EST\sub{n.left, t\sub{i}$^\prime$})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{i}$^\prime$})} $\in$ \nf{ops(EST\sub{n, t\sub{t}$^\prime$})} are in the block returned by \nf{n.CreateBlock()} by lemma 10. 
%By Lemma~\ref{lem::createBlock} \nf{new} contains \nf{n}'s childrens' established blocks after Line \ref{readHead} which is  appended to \nf{n.blocks[]} by \nf{TryAppend} in Line 313.
\end{proof}
\begin{figure}[hbt]
  \center\includegraphics[width=5in]{pics/trueRefresh}
  \caption{New established operations of the left child are in the new block.}
\end{figure}
\begin{lemma}[Stronger True Refresh] \label{lem::prectrueRefresh}
  Let $t_i$ be the time an instance of \nf{n.Refresh()} read the head (Line \ref{readHead}) and $t_t$ be the time its \nf{TryAppend(new, s)} terminates with  and returns \nf{true} (Line \ref{cas}). We have \nf{ops(EST\sub{n.left, t\sub{i}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{i}})} $\subseteq$ \nf{ops(n.blocks)}.
\end{lemma}
\begin{definition}
  An instance of \nf{Refresh()} is successful iff its \nf{TryAppend(new, s)} terminates with  and returns \nf{true}.
\end{definition}
%\it{Mention this new block is establish in t\sub{t}. last sentece is not complete}

@@

\begin{definition}
Let $_{R_1}t$ be the time $R_1$ is invoked and $t_{R_2}$ be the time $R_2$ terminates.
  $_{line}t$ is the immediate time before running Line $line$.   $t_{line}$ is the immediate time after running Line $line$. $_{line}t^{op}$ is the immediate time before running Line $line$ of operation $op$.   $t_{line}^{op}$ is the immediate time after running Line $line$ of operation $op$.
\end{definition}

\begin{lemma}[Double Refresh] \label{doubleRefresh}
  Consider two consecutive instances $R_1$, $R_2$ of \nf{Refresh()} on internal node \nf{n} by a process $p$. If $R_1$ and $R_2$ both fail and return false, then we have \nf{ops(EST\sub{n.left, \sub{R1}t})} $\cup$ \nf{ops(EST\sub{n.right, \sub{R1}t})} $\subseteq$ \nf{ops(EST\sub{n, t\sub{R2}})}.
\end{lemma}
\begin{proof}

If $R_2$ reads some value greater than $i+1$ in Line \ref{readHead} it means a successful instance of \nf{Refresh()} performed its Line \ref{readHead} after $t_{\ref{readHead}}^{R_1}$ and finished its Line \ref{incrementHead1} or \ref{incrementHead2} before $t_{\ref{readHead}}^{R_2}$, from Lemma~\ref{lem::prectrueRefresh} by the end of this instance \nf{ops(EST\sub{n.left, t\sub{1}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{1}})} has been propagated.

 Let $R_1$ read $i$ and $R_2$ read $i+1$ from Line \ref{readHead}. As $R_2$'s \nf{TryAppend()} returns \nf{false}, there is another successful instance $R_2^\prime$ of \nf{n.Refresh()} that has done \nf{TryAppend()} successfully into \nf{n.blocks[i+1]} before $R_2$ tries to append. Since $R_2^\prime$ creates the block after reading the value $i+1$ from \nf{n.head} (Line \ref{readHead}) and $R_1$ reads the value $i$ from \nf{n.head} and the \nf{head}'s value is increasing by Lemma \ref{lem::headProgress} then $t_{R2^\prime \; \ref{readHead}} > t_{R1 \; \ref{readHead}} > _{R_1}t$ (see Figure \ref{fig::doubleRefresh}). By Lemma \ref{lem::prectrueRefresh} after $R_2^\prime$'s \nf{CAS} ($t_{\ref{cas}}^{R_2^\prime}$) we have \nf{ops(EST\sub{n.left, t\sub{1}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{1}})} $\subseteq$ \nf{ops(n.blocks)}. Also by Lemma \ref{lem::headInc} on $R_2$ the value  of \nf{n.head}  head is more than $i+1$ after $R_2^\prime$ terminates, so the block appended by $R_2^\prime$ to \nf{n} is established by then ($\nf{n.head}\geq i+2 > i+1$). To summarize, $_{R_1}t$ is before $R_2^\prime$'s read of \nf{n.head} ($t_{\ref{readHead}}^{R_2^\prime}$) and $R_2^\prime$'s successful \nf{CAS} is before $R_2$'s termination. So, by Lemma \ref{lem::prectrueRefresh}, \nf{ops(EST\sub{n.left, t\sub{1}})} $\cup$ \nf{ops(EST\sub{n.right, t\sub{1}})} $\subseteq$ \nf{ops(EST\sub{n, t\sub{2}})}.
%It is obvious that the \nf{new} constructed by the second Refresh in Line 36 contains the blocks in \nf{new} by $R_1$ which $R_1^\prime$ did not contain, since \nf{n.head} is only increasing (Maybe Lemma~\ref{lem::oldnewOrder}). If $R_2$ succeeds by Lemma~\ref{lem::trueRefresh} the claim holds. If not, it is deduced that \nf{n.blocks[h]} was not \nf{null} before $R_2$'s \nf{CAS}. Furthermore, \nf{n.blocks[h]} was \nf{null} before reading \nf{h} by $R_1$. So there is a successful \nf{Refresh()} after the read of \nf{h} in $R_1$ and before the \nf{CAS} of $R_2$. This \nf{Refresh()} contains all the new established operations before Line 35 (Maybe Lemma~\ref{lem::oldnewOrder}) and by Lemma~\ref{lem::trueRefresh} our claim holds. See Figure~\ref{}.
\end{proof}

\begin{figure}[hbt]
  \center\includegraphics[width=6in]{pics/compactdouble.png}
  \caption{\label{fig::doubleRefresh}$_{R_1}t<$ $t_{\ref{readHead}}^{R_1}$ $<$ incrementing \nf{n.head} from $i$ to $i+1$ $<$ $t_{\ref{readHead}}^{R_2^\prime}$ $<$ $t_{\ref{cas}}^{R_2^\prime}$ $<$ incrementing \nf{n.head} from $i+1$ to $i+2$ $<t_{R_2}$} 
\end{figure}

@@

\begin{corollary}
  \nf{ops(EST\sub{n.left, \sub{\ref{firstRefresh}}t})} $\cup$ \nf{ops(EST\sub{n.right, \sub{\ref{firstRefresh}}t})} $\subseteq$ \nf{ops(EST\sub{n, t\sub{\ref{secondRefresh}})}}
\end{corollary}
\begin{proof}
  If the first \nf{Refresh()} in line \ref{firstRefresh} returns \nf{true} then by Lemma \ref{lem::trueRefresh} the claim holds. Also if first \nf{Refresh()} failed and the second \nf{Refresh()} succeeded the claim still holds by Lemma \ref{lem::trueRefresh}. Finally, if both failed the claim is satisfied by Lemma \ref{doubleRefresh}.
\end{proof}

@@

\begin{lemma}[Get correctness] \label{get}
If $\nf{n.blocks[b].num\sub{enq}} \geq \nf{i}$ then \nf{n.GetENQ(b,i)} returns the \nf{element} enqueued by $E_i(n,b)$.
\end{lemma}
\begin{proof}
We are going to prove this lemma by induction on the height of node \nf{n}. For the base case, \nf{n} is a leaf. Leaf blocks each contain exactly one operation, so by the hypothesis, only \nf{n.GetENQ(b,1)} can be called and only when \nf{n.blocks[b]} contains an enqueue. At Line \ref{getBaseCase}, \nf{n.GetENQ(b,1)} returns the \nf{element} of the \nf{enqueue} operation stored in the $b$th block of leaf \nf{n} .

For the induction step we prove \nf{n.GetENQ(b, i)} returns $E_i(n,b)$, assuming \nf{n.child.GetENQ(subblock, i)} returns $E_i(n.child,b)$. We argue that  Line \ref{leftOrRight} correctly decides whether the \nf{i}th enqueue in  \nf{b}th block  of internal node \nf{n} is in the  left child or right child subblocks of \nf{n.blocks[b]}. From Definition  \ref{ordering} of $E(n,b)$ we know enqueue operations in a block are ordered from left to right and since the leaves of the tree are ordered by process id from left to right, thus operations from the left subblocks come before operations from the right subblocks in a block (See Figure \ref{figGet}). Furthermore the \nf{num\sub{enq-left}} field  in \nf{n.blocks[b]} stores the number of \nf{enqueue()} operations from the blocks's subblocks in the left child of \nf{n}. So the $i$th enqueue operation is propagated from the right child if \nf{i} is greater than \nf{b.num\sub{enq-left}}. Otherwise we should search for the $i$th enqueue in the left child. By definition \ref{def::subblock} and \ref{def::ops} we need to search in subblocks of \nf{n.blocks[b]} from the range \texttt{n.left.blocks[n.blocks[i-1].end\textsubscript{left}+1..n.blocks[i].end\textsubscript{left}]} $\cup$ \texttt{n.right.blocks[n.blocks[i-1].end\textsubscript{right}+1..n.blocks[i].end\textsubscript{right}]}.

If the $i$th \nf{enqueue} of \nf{n.blocks[b]} is in the left child it would be $i$th enqueue in \texttt{n.left.blocks[n.blocks[i-1].end\textsubscript{left}+1..n.blocks[i].end\textsubscript{left}]} by Definition \ref{def::subblock}. Also, we know there are $eb=n.blocks[b-1].sum_{enq-left}$ enqueues in the blocks before this range, so $E_i(n,b)$ is $E_{i+eb}(n.left)$ which is $E_{i^\prime}(n.left,b^\prime)$ for some $b^\prime$ and $i^\prime$. We can compute $b^\prime$ and then search for $i+eb$th enqueue in \tt{n.left}, where $i^\prime$ is \tt{i+eb-n.left.blocks[}$b^\prime-1$\tt{].sum\sub{enq}}. The parameters in Line \ref{leftChildGet} are for searching $E_{i+eb}(n.left)$ in \nf{n.left.block} in the expected range of blocks, so this \nf{BSearch} returns the index of the subblock containing $E_i(n,b)$.

Otherwise the enqueue we are looking for is in the right child. Then, there are \nf{n.blocks[b].num\sub{enq-left}} enqueues ahead of it in \nf{n.blocks[b]} but not in \texttt{n.right.blocks[n.blocks[i-1].end\textsubscript{right}+1..n.blocks[i].end\textsubscript{right}]}. So we need to search for \nf{i-n.blocks[b].num\sub{enq-left}+ n.blocks[b-1].sum\sub{enq-right}} (Line \ref{rightChildGet}). Other parameters for the left child are chosen similarly to the way they were chosen for the right child.

So, in both cases the direct subblock containing $E_i(n,b)$ is computed in Lines \ref{leftChildGet} and \ref{rightChildGet}.
 Finally, \nf{n.child.GetENQ(subblock, i)} is invoked on the subblock containing $E_i(n,b)$ and it returns $E_i(n,b)$ by the hypothesis of the induction.
\end{proof}

@@

\begin{lemma}[DSearch correctness] \label{dsearch}
Assume $\nf{root.blocks[end].sum\sub{enq}}\geq\nf{e}$ and $E_e(root)$'s \nf{element} is the response to some \nf{Dequeue()} operation in \nf{root.blocks[end]}.  \nf{DSearch(e, end)} returns \nf{<b, i>} such that $E_i(root,b)= E_e(root)$.
\end{lemma}
\begin{proof}
It is trivial to see that the doubling search from \nf{root.blocks[end]} to \nf{root.blocks[0]} will find $E_e(root)$ eventually. Because \nf{root.blocks[].sum\sub{enq}} is an increasing value from 0 to some value greater than \nf{e}. So there is a \nf{b} that $\nf{root.blocks[b].sum\sub{enq}}>e$ but $\nf{root.blocks[b-1].sum\sub{enq}}<e$.

First we show $\nf{end}-\nf{b} \leq 2 \times \big(\nf{root.blocks[b].\size}+\nf{root.blocks[end].\size+1\big)$. From line \ref{addOP}, we know that size of the every block in the tree is greater than 0. So each block in \nf{root.blocks[b..end]} contains at least one \nf{Enqueue} or at least one \nf{Dequeue}. Suppose there were more than \nf{root.blocks[b].\size}\nf{Dequeue}s in \nf{root.blocks[b+1..end-1]}. Then the  queue would become empty at some point after \nf{blocks[b]}'s last operations and before \nf{root.blocks[end]}'s first operation. Which means the response to to a \nf{Dequeue} in \nf{root.blocks[end]} could not be in $E(n,b)$. Furthermore since the size of the queue would become \nf{root.blocks[end].\size} after the \nf{root.blocks[end]}, there cannot be more than \nf{root.blocks[b].\size + root.blocks[end].\size} \nf{Enqueue}s. Because there can be at most \nf{root.blocks[b].\size}\nf{Dequeue}s and the final size is \nf{root.blocks[end].\size}. Overall there can be at most $2 \times$\nf{root.blocks[b].\size+ root.blocks[end].\size} operations in \nf{root.blocks[b+1..end-1]} and since each block size is $\geq 1$ thus there are at most $2 \times$\nf{root.blocks[b].\size+ root.blocks[end].\size} blocks in between \nf{root.blocks[b]} and \nf{root.blocks[end]}. So \nf{end-b}$\leq 2 \times$\nf{root.blocks[b].\size+root.blocks[end].\size+1}. See Figure \ref{fig::doubling}.

@@

\begin{lemma}[Index correctness]
 If $\nf{n.blocks[b].num\sub{deq}}\geq\nf{i}$ then \nf{n.IndexDEQ(b,i)} returns the rank in $D(root)$ of $D_{n,b,i}$.
\end{lemma}
\begin{proof}
We will prove this by induction on the distance of \nf{n} from the \nf{root}. We can see the base case where \nf{n} is root is trivial (Line \ref{indexBaseCase}).
  In the non-root nodes \nf{n.IndexDEQ(b,i)} computes the superblock of the $i$th Dequeue in the $b$th block of \nf{n} in \nf{n.parent} by Lemma \ref{superBlock} (Line \ref{computeSuper}). After that the order in $D(n.parent, superblock)$ is computed. Note that by Lemma \ref{blockSize} in each block there is at most one operation from each process and operations of one type are ordered based on the order in the subblocks (See Figure \ref{fig::orderFromSubblocks}). Finally \nf{index()} is called on \nf{n.parent} recursively and it returns the correct response from induction hypothesis. If the operation was propagated from the right child the number of dequeues from the left child are added to it (Line \ref{considerRight}), because the left child operations come before the right child operations (Definition \ref{ordering}).
\end{proof}
\textit{Make sure to show preconditions of all invocation of \nf{BSearch} are satisfied.}
